From Real-time Attention Assessment to ``With-me-ness'' in Human-Robot Interaction

### 2015-09-22 paper.tex
Various minor changes

We then propose to introduce the concept of ``with-me-ness'', borrowed from the
field of {\it Computer-Supported Collaborative Learning}, as an
\emph{in-the-moment} measurement of the quality of interaction. We argue that
this new metric, more specific than the broad concept of engagement, may be well
suited for operational, real-time assessment of an on-going interaction.

@@

Looking at techniques that both operate on-line and have been deployed in field
experiments, one finds that most approaches rely on head pose estimation alone
(no eye gaze tracking) and are most often based on depth sensors (RGB-D) .
Fanelli~\etal provides a overview of these approaches in~\cite{fanelli2012real},
and recent examples include~\cite{baxter2014tracking} or~\cite{anzalone}.

@@

The width of the human field of view has been discussed in the literature.
Holmqvist~\cite{holmqvist2011eye} specifies that the visual human
range is $ \pm  40\degree $ in the
horizontal and $ \pm 25\degree $ in the vertical.
In~\cite{walker1980clinical}, Spector provides a more detailed specification
for each eye splitting the vertical range into $ 60\degree $ the upper region
and $ 75\degree $ the lower one. Moreover, the horizontal range gets separated
in $ 60\degree $ inwards (towards the nose) and $ 95\degree $ outwards.
Previous work on visual perspective taking for social robotics
application~\cite{sisbot2011situation} propose to use a cone of 30$\degree$ of
aperture. In the present implementation, the first approach explained has been
chosen representing the field of view using a cone with an elliptical cross section 
and the dimensions stated by Holmqvist.

We then approximate the visual focus of attention (VFoA) of the human to the
objects which lie inside the field of attention (Figure~\ref{fig:vfoa}). At a
given time, more than one object can therefore be \emph{in focus}.

@@

Six children (ages 5 to 6, 3 boys, 3 girls) were enrolled for this study.
The study took place at school, in a separate room (the computer
lab). The participants were chosen by the teacher, and would come one after the
other for a face-to-face interaction with a {\sc Nao} robot (average duration:
19.6 min, $SD=1.58$).

@@

The experiment was carried with an Aldebaran's {\sc Nao} robot, using ROS as a
middleware to build the attention estimation pipeline (Figure~\ref{fig:system}).
Head pose estimation, presented in section~\ref{sec:vfoa}, builds on the {\tt
dlib} and OpenCV libraries; the pose transformations are handled by the ROS {\sc
tf} library. The implementation is open-source and available at
\url{https://github.com/chili-epfl/attention-tracker}.

>>>>

We secondly propose to introduce the concept of ``with-me-ness'', borrowed from the
field of {\it Computer-Supported Collaborative Learning}, as an
\emph{in-the-moment} measurement of the quality of interaction. We argue that
this new metric, more specific than the broad concept of engagement, may be well
suited for operational, real-time assessment of an on-going interaction.

@@

Looking at techniques that both operate on-line and have been deployed in field
experiments, one finds that most approaches rely on head pose estimation alone
(no eye gaze tracking) and are most often based on depth sensors (RGB-D) .
Fanelli~\etal provides a overview of these approaches in~\cite{fanelli2012real},
and recent examples include~\cite{baxter2014tracking, anzalone}.

@@

The dimensions of the human field of view have been measured at different level
of accuracy. Holmqvist~\cite{holmqvist2011eye} models it with an horizontal
aperture of $ \pm 40\degree $ and a vertical aperture of $ \pm 25\degree $,
while Spector~\cite{walker1980clinical} for instance suggests $60\degree$ up,
$75\degree$ down, $60\degree$ inwards (towards the nose) and $95\degree$
outwards.  Previous work on visual perspective taking for social
robotics~\cite{sisbot2011situation} simply use a cone of $30\degree$.  In this
work, we follow Holmqvist and thus model the field of view as a cone with an elliptical
cross-section of $80\degree$ by $50\degree$.

We then approximate the visual focus of attention (VFoA) of the human to the
objects which lay inside this field of view (Figure~\ref{fig:vfoa}). At a
given time, more than one object can therefore be \emph{in focus}.

@@

Six children (ages 5 to 6, 3 boys, 3 girls) were enrolled for this study.
The study took place at school, in a separate room (the computer
lab). The participants were chosen by the teacher, and would come one after the
other for a face-to-face interaction with a {\sc Nao} robot ($M=19.6$ min, $SD=1.58$).

@@

The experiment was carried with an Aldebaran's {\sc Nao} robot, using ROS as a
middleware to build the attention estimation pipeline (Figure~\ref{fig:system}).
Head pose estimation, presented in section~\ref{sec:vfoa}, builds on the {\tt
dlib} and OpenCV libraries; the pose transformations are handled by the ROS {\sc
tf} library. The same {\sc tf} library is used to represent as individual frames
the possible point of interests: an object is considered to be in focus when its
frame lays within the field of attention of the participant
(Figure~\ref{fig:vfoa}).  The implementation is open-source and available at
\url{https://github.com/chili-epfl/attention-tracker}.
